---
title: "Solo Project"
format: html

---

# Project Overview

This project visualizes the global diversity of cheeses by country using a bubble map. The size and color of each circle represent the number of unique cheese varieties attributed to that country.

## Goal

Apply effective visualization principles:
 - Simplify gridlines and distractions
 - Use perceptually uniform color scale
 - Clear labels and consistent fonts
 - Provide attribution and interpretation

```{r}
library(tidyverse)
library(readr)
library(maps)
cheeses <- read_csv("cheeses.csv")
coords <- read_csv("world_country_and_usa_states_latitude_and_longitude_values.csv")

cheese_country <- cheeses %>%
  count(country, name = "cheese_count")

cheese_map_data <- left_join(cheese_country, coords, by = "country") %>%
  filter(!is.na(latitude), !is.na(longitude))
```

## World Cheese Map

```{r}
ggplot(cheese_map_data, aes(x = longitude, y = latitude)) +
  borders("world", colour = "gray60", fill = "gray98") +
  geom_point(aes(size = cheese_count, color = cheese_count), alpha = 0.85) +
  scale_color_viridis_c(option = "plasma", name = "Cheese Count") +
  scale_size_continuous(range = c(2, 12)) +
  labs(
    title = "🧀 Cheese Diversity Around the World",
    subtitle = "Each circle shows the number of unique cheeses per country",
    x = NULL, y = NULL,
    caption = "Data: cheeses.csv + country coordinates • Author: Isa"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(face = "bold", size = 20),
    plot.subtitle = element_text(size = 14),
    panel.grid = element_blank(),
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    plot.caption = element_text(margin = margin(t = 10), size = 10)
  )
```

##  Interpretation

Cheese is more than a food — it’s a cultural expression tied to geography, tradition, and climate. This map illustrates where in the world cheese has become a central part of cuisine and identity.

Countries like **France** and **Italy** appear with large, colorful bubbles, reflecting their deep histories in dairy and fermentation. These regions produce hundreds of distinct varieties, shaped by regional environments and centuries-old recipes.

Meanwhile, other countries like **Switzerland**, **United Kingdom**, and **Greece** also stand out, offering famous cheeses like Emmental, Cheddar, and Feta — each with its own story and significance.

Even countries with fewer known varieties may represent untapped or undocumented traditions. This visualization invites curiosity: what makes a region specialize in cheese? What is missing in areas with little data?

Through visualization, we not only quantify cheese diversity but spark appreciation for how food connects people to place.

##  Data Sources

- `cheeses.csv`: dataset of world cheeses
- `world_country_and_usa_states_latitude_and_longitude_values.csv`: country coordinates
